{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-07-19 sftrends_manu.Rnw
second case study figs and tabs done

%acronyms
\acrodef{chla}[chl-\textit{a}]{chlorophyll \textit{a}}
\acrodef{din}[DIN]{dissolved inorganic nitrogen}
\acrodef{sfe}[SFE]{San Francisco Estuary}
\acrodef{wrtds}[WRTDS]{Weighted Regressions on Time, Discharge, and Season}

@@

@@

@@

# globals
ylab1 <- expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-', ' (mg ', L^-1, ')'))
ylab2 <- expression(paste(NH [4] ^ '2+', ' (mg ', L^-1, ')'))
flolab <- expression(paste('ln-Flow (', m^3, ' ', s^-1, ')'))
ptheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(),
    axis.line.y = element_line()
  )
col_vec <- RColorBrewer::brewer.pal(9, 'Blues')[c(9, 8, 7)]
dynacol_vec <- RColorBrewer::brewer.pal(9, 'Spectral')
xlims <- range(nomod$date)

@@

# colors
cols <- RColorBrewer::brewer.pal(9, 'Set1')[1:3]

## first set of plots
pdin1 <- prdnrmplot(toplo1$mod[[2]], col_vec = cols[2], logspace = F, min_mo = 10) +
  ptheme + 
  scale_y_continuous(ylab1) +
  theme(legend.position = 'top', legend.title = element_blank())
pleg1 <- g_legend(pdin1)
pdin1 <- pdin1 + theme(legend.position = 'none')

@@

pchl1 <- prdnrmplot(toplo1$mod[[1]], col_vec = cols[2], logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none') + 
  scale_y_continuous(ylab2)

@@

psio1 <- prdnrmplot(toplo1$mod[[3]], col_vec = cols[2], logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text()) +
  scale_x_date('Year') +
  scale_y_continuous(ylab3)

@@

## second set of plots
pdin2 <- prdnrmplot(toplo2$mod[[2]], col_vec = cols[2], logspace = F, min_mo = 10) +
  ptheme + 
  scale_y_continuous(ylab1) +
  theme(legend.position = 'none')

@@

pchl2 <- prdnrmplot(toplo2$mod[[1]], col_vec = cols[2], logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none') + 
  scale_y_continuous(ylab2)

@@

psio2 <- prdnrmplot(toplo2$mod[[3]], col_vec = cols[2], logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text()) +
  scale_x_date('Year') +
  scale_y_continuous(ylab3)

@@

@@

mobrks <- c(-Inf, 3, 6, 9, Inf)
yrbrks <- c(-Inf, 1985, 1994, 2003, Inf)
molabs <- c('JFM', 'AMJ', 'JAS', 'OND')
yrlabs <- c('1976-1985', '1986-1994', '1995-2003', '2004-2013')

# summarize din, sio2, chl trends
din <- wrtdstrnd(dinraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
din <- wrtdstrnd(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

@@

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.'

>>>>

%acronyms
\acrodef{chla}[chl-\textit{a}]{chlorophyll \textit{a}}
\acrodef{din}[DIN]{dissolved inorganic nitrogen}
\acrodef{sfe}[SFE]{San Francisco Estuary}
\acrodef{wrtds}[WRTDS]{Weighted Regressions on Time, Discharge, and Season}
\acrodef{wwtp}[WWTP]{wastewater treatment plant}

@@

What was reduction in ammonium starting in 2002?

@@

Trends as percent change depend on the mean value, lower values will have larger percent changes.

@@

# globals
ylab1 <- expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-', ' (mg ', L^-1, ')'))
ylab2 <- expression(paste(NH [4] ^ '+', ' (mg ', L^-1, ')'))
flolab <- expression(paste('ln-Flow (', m^3, ' ', s^-1, ')'))
ptheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(),
    axis.line.y = element_line()
  )
col_vec <- RColorBrewer::brewer.pal(9, 'Blues')[c(9, 8, 7)]
dynacol_vec <- RColorBrewer::brewer.pal(9, 'Spectral')
xlims <- range(nomod$date)

@@

# colors
cols <- RColorBrewer::brewer.pal(9, 'Set1')[1:3]
col_vec <- RColorBrewer::brewer.pal(9, 'Blues')[c(9, 8, 7)]

## first set of plots
pdin1 <- prdnrmplot(toplo1$mod[[2]], col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  scale_y_continuous(ylab1) +
  theme(legend.position = 'top', legend.title = element_blank())
pleg1 <- g_legend(pdin1)
pdin1 <- pdin1 + theme(legend.position = 'none')

@@

pchl1 <- prdnrmplot(toplo1$mod[[1]], col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none') + 
  scale_y_continuous(ylab2)

@@

psio1 <- prdnrmplot(toplo1$mod[[3]], col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text()) +
  scale_x_date('Year') +
  scale_y_continuous(ylab3)

@@

## second set of plots
pdin2 <- prdnrmplot(toplo2$mod[[2]], col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  scale_y_continuous(ylab1) +
  theme(legend.position = 'none')

@@

pchl2 <- prdnrmplot(toplo2$mod[[1]], col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none') + 
  scale_y_continuous(ylab2)

@@

psio2 <- prdnrmplot(toplo2$mod[[3]], col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text()) +
  scale_x_date('Year') +
  scale_y_continuous(ylab3)

@@

% trends in no23, nh at p8
<<p8chg>>=

data(h2dat)
dt <- as.Date('2007-05-01')

noraw_bf <- filter(h2dat, resvar %in% 'no23') %>% 
  .$mod %>% 
  .[[1]] %>% 
  .[which(.$date < dt), ]
noraw_af <- filter(h2dat, resvar %in% 'no23') %>% 
  .$mod %>% 
  .[[1]] %>% 
  .[which(.$date >= dt), ]
nhraw_bf <- filter(h2dat, resvar %in% 'nh') %>% 
  .$mod %>% 
  .[[1]] %>% 
  .[which(.$date < dt), ]
nhraw_af <- filter(h2dat, resvar %in% 'nh') %>% 
  .$mod %>% 
  .[[1]] %>% 
  .[which(.$date >= dt), ]

mobrks <- c(0, 3, 6, 9, 12)
yrbrks_bf <- c(1975, 2007)
yrbrks_af <- c(2006, 2013)
molabs <- c('JFM', 'AMJ', 'JAS', 'OND')
yrlabs_bf <- '1976-2007'
yrlabs_af <- '2008-2013'
faclevs <- c('1976-2007_before', '2008-2013_after', paste0(molabs, '_before'), paste0(molabs, '_after'))

# summarize din, sio2, chl trends
no_bf <- wrtdstrnd(noraw_bf, mobrks, yrbrks_bf, molabs, yrlabs_bf, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'before', var = 'no')
no_af <- wrtdstrnd(noraw_af, mobrks, yrbrks_af, molabs, yrlabs_af, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'after', var = 'no')
nh_bf <- wrtdstrnd(nhraw_bf, mobrks, yrbrks_bf, molabs, yrlabs_bf, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'before', var = 'nh')
nh_af <- wrtdstrnd(nhraw_af, mobrks, yrbrks_af, molabs, yrlabs_af, aves = TRUE, min_mo = 10) %>% 
  mutate(period = 'after', var = 'nh')

# combine for trends for table
totab <- rbind(no_bf, no_af, nh_bf, nh_af) %>% 
  mutate(
    chg = round(chg, 1), 
    chg = ifelse(chg < 0, chg, paste0('\\textit{\\textbf{', chg , '}}')), 
    ave = as.character(round(ave, 2))
    ) %>% 
  gather('est', 'val', chg:ave) %>% 
  unite('col', var, est) %>% 
  spread(col, val) %>% 
  unite('period', cat, period) %>% 
  mutate(period = factor(period, levels = faclevs, labels = faclevs)) %>% 
  arrange(period) %>% 
  mutate(cat = c(yrlabs_bf, yrlabs_af, molabs, molabs)) %>% 
  data.frame(sep = c(rep('Annual', 2), rep('Seasonal, pre', 4), rep('Seasonal, post', 4)), .) %>% 
  select(sep, cat, no_ave, no_chg, nh_ave, nh_chg)

# final table formatting
segs <- unique(totab$sep)
cats <- totab$cat
totab <- totab[,-c(1:2)]
vars <- c('NO$_{2}^{-}$/NO$_{3}^{2-}$', 'NH$_{4}^{+}$')
names(totab) <- rep(c('Mean', '\\% change'), 2)

cap.val<-'Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \\ac{wwtp} upgrades upstream of station P8. Upgrades were completed in May 2007 at the City of Tracy \\ac{wwtp} (San Joaquin County, \\cref{fig:tracy}).  Summaries are means and percent changes based on annual means within the pre- and post-upgrade time periods (1976-2007, 2008-2013).  Increasing values are in bold-italics.'

latex( 
  totab,
  file = '',
  rowlabel = 'Period',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = segs,
  n.rgroup = c(2, rep(4, 2)),
  cgroup = vars,
  n.cgroup = c(2, 2),
  rowname = cats,
  label = 'tab:p8chg'
  )

@@

mobrks <- c(0, 3, 6, 9, 12)
yrbrks <- c(1975, 1985, 1994, 2003, 2013)
molabs <- c('JFM', 'AMJ', 'JAS', 'OND')
yrlabs <- c('1976-1985', '1986-1994', '1995-2003', '2004-2013')

# summarize din, sio2, chl trends
din <- wrtdstrnd(dinraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
din <- wrtdstrnd(dinraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., din) %>% 
  data.frame(res = 'din', .)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
chl <- wrtdstrnd(chlraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., chl) %>% 
  data.frame(res = 'chl', .)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks, molabs, yrlabs, aves = TRUE, min_mo = 10)
sio <- wrtdstrnd(sioraw, mobrks, yrbrks = c(-Inf, Inf), molabs, yrlabs = c('1976-2013'), aves = TRUE, min_mo = 10) %>% 
  .[1, ] %>% 
  rbind(., sio) %>% 
  data.frame(res = 'sio', .)

@@

cap.val<-'Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.'

### 2016-07-19 sftrends_manu.tex
second case study figs and tabs done

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Tue Jul 19 09:56:38 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% trends in din, chl, sio2 at d7
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:d7chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:d7chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&0.4&-4.2&&4.6&-82.4&&12.5&\textit{\textbf{30}}\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&0.4&-15.2&&10.2&-67.9&&10.7&\textit{\textbf{8.4}}\tabularnewline
~~1986-1994&0.4&\textit{\textbf{15.4}}&&3.3&-48.8&&12.3&\textit{\textbf{13.5}}\tabularnewline
~~1995-2003&0.5&\textit{\textbf{4.3}}&&2.1&-3.9&&13.5&\textit{\textbf{3.7}}\tabularnewline
~~2004-2013&0.5&-10.6&&2.5&\textit{\textbf{31.7}}&&13.7&-1.8\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&0.5&-4.9&&2.7&-56&&15&\textit{\textbf{4.4}}\tabularnewline
~~AMJ&0.4&\textit{\textbf{17.4}}&&6.1&-74.3&&12.9&\textit{\textbf{20.9}}\tabularnewline
~~JAS&0.3&\textit{\textbf{40.6}}&&6.2&-89&&10.9&\textit{\textbf{79.1}}\tabularnewline
~~OND&0.5&\textit{\textbf{13.1}}&&2.7&-83.8&&11.7&\textit{\textbf{33.7}}\tabularnewline
\hline
\end{tabular}\end{center}

@@

% trends in din, chl, sio2 at c10
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:c10chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:c10chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&1.3&-16.9&&27.6&-4.9&&14.5&-23.1\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&1.3&\textit{\textbf{16.6}}&&26.6&-20.5&&15.7&-5\tabularnewline
~~1986-1994&1.4&\textit{\textbf{6.6}}&&27.5&\textit{\textbf{23.7}}&&14.8&-3.9\tabularnewline
~~1995-2003&1.4&-1.1&&29.3&-15.2&&14.1&-5.1\tabularnewline
~~2004-2013&1.2&-32.6&&27.3&\textit{\textbf{11}}&&13&-9.6\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&1.4&-20.5&&12&\textit{\textbf{14.8}}&&14.3&-16.9\tabularnewline
~~AMJ&1&-26.4&&43&\textit{\textbf{3.1}}&&13.1&-27.5\tabularnewline
~~JAS&1.3&-3.7&&46.6&-13.6&&14.8&-26.1\tabularnewline
~~OND&1.6&-6.8&&8.7&-18.5&&15.6&-16\tabularnewline
\hline
\end{tabular}\end{center}

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Tue Jul 19 14:18:21 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% trends in no23, nh at p8
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(2, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2), rowname = cats,     label = "tab:p8chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in nitrite/nitrate and ammonium (mg L$^{-1}$) concentrations before and after \ac{wwtp} upgrades upstream of station P8. Upgrades were completed in May 2007 at the City of Tracy \ac{wwtp} (San Joaquin County, \cref{fig:tracy}).  Summaries are means and percent changes based on annual means within the pre- and post-upgrade time periods (1976-2007, 2008-2013).  Increasing values are in bold-italics.\label{tab:p8chg}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries NO$_{2}^{-}$/NO$_{3}^{2-}$}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NH$_{4}^{+}$}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries Annual}&&&&&\tabularnewline
~~1976-2007&1.27&\textit{\textbf{75.7}}&&0.33&\textit{\textbf{71.3}}\tabularnewline
~~2008-2013&1.47&-6&&0.11&-62.6\tabularnewline
\hline
{\bfseries Seasonal, pre}&&&&&\tabularnewline
~~JFM&1.46&\textit{\textbf{77.7}}&&0.49&\textit{\textbf{57.1}}\tabularnewline
~~AMJ&1&\textit{\textbf{75.3}}&&0.13&\textit{\textbf{63.4}}\tabularnewline
~~JAS&1.08&\textit{\textbf{97.9}}&&0.13&\textit{\textbf{151.9}}\tabularnewline
~~OND&1.53&\textit{\textbf{81.9}}&&0.54&\textit{\textbf{102.4}}\tabularnewline
\hline
{\bfseries Seasonal, post}&&&&&\tabularnewline
~~JFM&1.76&-3.4&&0.14&-47\tabularnewline
~~AMJ&0.94&-5.5&&0.07&-45.3\tabularnewline
~~JAS&1.17&-10&&0.07&-50.3\tabularnewline
~~OND&1.99&-4.6&&0.15&-58.5\tabularnewline
\hline
\end{tabular}\end{center}

@@

% trends in din, chl, sio2 at d7
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:d7chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg L$^{-1}$), chlorophyll ($\mu$g L$^{-1}$), and silicon dioxide (mg L$^{-1}$) concentrations for different time periods at station D7. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:d7chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&0.4&\textit{\textbf{27.8}}&&4.4&-81.6&&12.6&\textit{\textbf{30.2}}\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&0.4&\textit{\textbf{10.2}}&&9.8&-63.9&&10.8&\textit{\textbf{9}}\tabularnewline
~~1986-1994&0.4&\textit{\textbf{14.9}}&&3.4&-48.8&&12.2&\textit{\textbf{14.2}}\tabularnewline
~~1995-2003&0.5&\textit{\textbf{4}}&&2.1&-4.8&&13.5&\textit{\textbf{3.7}}\tabularnewline
~~2004-2013&0.5&-8.6&&2.5&\textit{\textbf{28.5}}&&13.7&-1.8\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&0.5&-4.9&&2.7&-56&&15&\textit{\textbf{4.4}}\tabularnewline
~~AMJ&0.4&\textit{\textbf{17.4}}&&6.1&-74.3&&12.9&\textit{\textbf{20.9}}\tabularnewline
~~JAS&0.3&\textit{\textbf{40.6}}&&6.2&-89&&10.9&\textit{\textbf{79.1}}\tabularnewline
~~OND&0.5&\textit{\textbf{13.1}}&&2.7&-83.8&&11.7&\textit{\textbf{33.7}}\tabularnewline
\hline
\end{tabular}\end{center}

% trends in din, chl, sio2 at c10
%latex.default(totab, file = "", rowlabel = "Period", caption = cap.val,     caption.loc = "top", rgroup = segs, n.rgroup = c(1, rep(4,         2)), cgroup = vars, n.cgroup = c(2, 2, 2), rowname = cats,     label = "tab:c10chg")%
\begin{table}[!tbp]
\caption{Summaries of flow-normalized trends in dissolved inorganic nitrogen (mg $L^{-1}$), chlorophyll ($\mu$g $L^{-1}$), and silicon dioxide (mg $L^{-1}$) concentrations for different time periods at station C10. Summaries are means and percent changes based on annual means within the time periods.  Increasing values are in bold-italics.\label{tab:c10chg}} 
\begin{center}
\begin{tabular}{lllcllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Period}&\multicolumn{2}{c}{\bfseries DIN}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries Chl-\textit{a}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries SiO$_2$}\tabularnewline
\cline{2-3} \cline{5-6} \cline{8-9}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{\% change}\tabularnewline
\hline
{\bfseries All}&&&&&&&&\tabularnewline
~~1976-2013&1.3&-16.5&&27.5&\textit{\textbf{9}}&&14.4&-21.7\tabularnewline
\hline
{\bfseries Annual}&&&&&&&&\tabularnewline
~~1976-1985&1.3&\textit{\textbf{12.1}}&&26.1&-3.3&&15.7&-4.3\tabularnewline
~~1986-1994&1.4&\textit{\textbf{6.3}}&&27.5&\textit{\textbf{23.5}}&&14.8&-4\tabularnewline
~~1995-2003&1.4&-0.8&&29.3&-15.3&&14.1&-5\tabularnewline
~~2004-2013&1.2&-29.4&&27.1&\textit{\textbf{5.1}}&&13.1&-8.7\tabularnewline
\hline
{\bfseries Seasonal}&&&&&&&&\tabularnewline
~~JFM&1.4&-20.5&&12&\textit{\textbf{14.8}}&&14.3&-16.9\tabularnewline
~~AMJ&1&-26.4&&43&\textit{\textbf{3.1}}&&13.1&-27.5\tabularnewline
~~JAS&1.3&-3.7&&46.6&-13.6&&14.8&-26.1\tabularnewline
~~OND&1.6&-6.8&&8.7&-18.5&&15.6&-16\tabularnewline
\hline
\end{tabular}\end{center}
